SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-03 docs/questioning.tex
Section 4 draft ready

A number of the frameworks define policies
based on information and features provided by
the Document Object Model ({\sc dom}) of a web page.
Specifically, developers must place all legitimate
scripts inside {\sc html} elements like {\tt div}.
The web browser (locus {\sc x}), parses the
{\sc dom} tree and executes scripts only when they
are contained in such elements. All other scripts
are treated according to the policies defined
on the server. Frameworks that support this
functionality are {\sc beep}~\cite{TNH07}
and {\sc dsi}~\cite{NSS06}. The main problem with
the above mechanisms is that they are vulnerable
to non-persistent {\sc xss} attacks that alter
the {\sc dom} of an already rendered
page~\cite{APKLM10}.
In addition, a recent variation of JavaScript
injection attacks known as return-to-JavaScript
attack~\cite{APKLM10} can be use to bypass
the above mechanisms. Such attacks share
many similarities with the {\it return-to-lib}
attack~\cite{SPPGMMB04}. When performing an
attack like this the malicious user transfers the
execution to points that contain script
code which in turn is used to perform the attack.

@@

\subsubsection{Runtime Tainting}

A runtime tainting scheme marks untrusted
(``tainted") data (for instance, a variable set
by a field in a web form) and traces its flow through
the program. If the variable is used in an expression
that sets another variable, that variable is also
marked as untrusted and so on. If any of these variables
is used to execute a potentially risky command (``sink") the
scheme may act accordingly.

Runtime tainting is provided as a feature in some
programming languages, such as Perl and Ruby.
By enabling the feature, Perl would refuse to run code
vulnerable to an {\sc sql} injection attack
(consider a tainted variable being used in a query)
and exit with an error message.

There are different implementations of these approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (locus {\sc z}).
However, this is a simplistic approach that may
lead to many false positives.
The following schemes apply further checks when
they identify that tainted data have reached a sink.
In particular, {\it Xu et al.}~\cite{XBS06}
tracks taint information at the level of bytes in memory.
To distinguish between legitimate and malicious uses
of untrusted data that reach a sink,
they search the data for suspicious symbols by using
regular expressions. {\sc csse}~\cite{PB05} associates
tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application
and other. When tainted data reaches a sink {\sc csse} performs
syntactic checks it based on the aforementioned metadata
(locus {\sc z}). {\it {\sc php} Aspis}~\cite{PMP11} works
in a similar way. To obtain metadata, it takes advantage
of specific features provided by {\sc php}.
Note that the above schemes detect potential attacks
in locus {\sc z}. This automatically means that they
cannot detect non-persistent JavaScript-driven
{\sc xss} attacks. {\sc wasc}~\cite{NLC07} 
though, can deal with such attacks because it
analyzes {\sc html} responses to check
if there are tainted data that contain scripts (locus {\sc y}).
A recent study showed that there are ways to circumvent
the majority of the above schemes~\cite{NBR14}.
Furthermore, most of them are hard to deploy since
the majority of input vectors, string operations
and output vectors must be instrumented.

@@

This category includes mechanisms that borrow
characteristics from more than one of the above categories.

@@

{\it Blueprint}~\cite{LV09} is a policy
enforcement mechanism that also utilizes parsed-trees.

{\it Diglossia}~\cite{SMS13} performs positive taint tracking and
parse-tree validation.

@@

	\multirow{7}{*}{Runtime Tainting}

>>>>

A number of the frameworks define policies
based on information and features provided by
the Document Object Model ({\sc dom}) of a web page.
Specifically, developers must place all legitimate
scripts inside {\sc html} elements like {\tt div}.
The web browser (locus {\sc x}), parses the
{\sc dom} tree and executes scripts only when they
are contained in such elements. All other scripts
are treated according to the policies defined
on the server. Frameworks that support this
functionality are {\sc beep}~\cite{TNH07}
and {\sc dsi}~\cite{NSS06}. The main problem with
the above mechanisms is that they are vulnerable
to non-persistent {\sc xss} attacks that alter
the {\sc dom} of an already rendered
page~\cite{APKLM10}.
In addition, a recent variation of JavaScript
injection attacks known as return-to-JavaScript
attack~\cite{APKLM10} can be used to bypass
the above mechanisms. Such attacks share
many similarities with the {\it return-to-lib}
attack~\cite{SPPGMMB04}. When performing an
attack like this the malicious user transfers the
execution to points that contain script
code which in turn is used to perform the attack.

@@

\subsubsection{Taint Tracking}
\label{sec:taint}

A taint tracking scheme marks untrusted
(``tainted") data (for instance, a variable set
by a field in a web form) and traces its flow through
the program. If the variable is used in an expression
that sets another variable, that variable is also
marked as untrusted and so on. If any of these variables
is used to execute a potentially risky command (``sink") the
scheme may act accordingly.

Taint tracking is provided as a feature in some
programming languages, such as Perl and Ruby.
By enabling the feature, Perl would refuse to run code
vulnerable to an {\sc sql} injection attack
(consider a tainted variable being used in a query)
and exit with an error message.

There are different implementations of these approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (locus {\sc z}).
However, this is a simplistic approach that may
lead to many false positives.
The following schemes apply further checks when
they identify that tainted data have reached a sink.
In particular, {\it Xu et al.}~\cite{XBS06}
tracks taint information at the level of bytes in memory.
To distinguish between legitimate and malicious uses
of untrusted data that reach a sink,
they search the data for suspicious symbols by using
regular expressions. {\sc csse}~\cite{PB05} associates
tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application
and other. When tainted data reaches a sink {\sc csse} performs
syntactic checks it based on the aforementioned metadata
(locus {\sc z}). {\it {\sc php} Aspis}~\cite{PMP11} works
in a similar way. To obtain metadata, it takes advantage
of specific features provided by {\sc php}.
Note that the above schemes detect potential attacks
in locus {\sc z}. This automatically means that they
cannot detect non-persistent JavaScript-driven
{\sc xss} attacks. Although they can be
modified to do so. {\sc wasc}~\cite{NLC07} 
though, can deal with such attacks because it
analyzes {\sc html} responses to check
if there are tainted data that contain scripts (locus {\sc y}).
A recent study showed that there are ways to circumvent
the majority of the above schemes~\cite{NBR14}.
Furthermore, most of them are hard to deploy since
the majority of input vectors, string operations
and output vectors must be instrumented.

@@

This category includes mechanisms that borrow
characteristics from more than one of the above categories.
Two of them focus on the detection of {\sc xss}
attacks and one of them focuses on {\sc dsl} code
injection attacks.

@@

{\it Blueprint}~\cite{LV09} is a policy
enforcement framework that also utilizes parsed-trees
to detect {\sc xss} attacks. To guarantee
that untrusted content is not executed, Blueprint generates
on the server-side a parse-tree from untrusted {\sc html}
to ensure that it does not contain any dynamic content.
Then the parsed-tree is transfered to the document
generator of the browser (locus {\sc x}) where untrusted
browser parsing behavior is ruled out.
Blueprint is an efficient countermeasure but it
imposes a significant overhead due to its
extensive parsing (see Table~\ref{tab:comp}).

{\it Diglossia}~\cite{SMS13} combines positive taint tracking
(see Subsubsection~\ref{sec:taint}) together with
parse-tree validation. Diglossia was based on the
theory of Ray and Ligatti~\cite{RL12b} (also see
Section~\ref{sec:attacks}) to detect {\sc dsl}
code injection attacks. In addition, it is actually
the first, and so far the only framework that detects
No{\sc sql} injection attacks.
When an application computes an output string (query),
Diglossia computes a ``shadow" of this string.
Specifically, it maps all characters introduced by
the application to a shadow character set.
This set does not contain any characters coming
from the tainted input. Then, the scheme
creates the parse-tree of the query that
is about to be executed with the parse-tree
of the ``shadow" and compares them (locus {\sc z}).
If the trees are not equal the application is probably
under attacks. Diglossia can be bypassed in the same
way as other taint tracking approaches~\cite{NBR14}.

@@

	\multirow{7}{*}{Taint Tracking}
